Predicting the effects of character displacement on food-web dynamics

### 2020-01-30 manuscript_ECD_model.Rmd
omitted personal identifying information for double-blind peer review at Am Nat

---
output:
  pdf_document:
    keep_tex: true
    toc: false
    number_sections: false
    df_print: paged
    fig_caption: true
    citation_package: natbib 
header-includes:
  - \usepackage{fullpage}
  - \linespread{1.5} 
  - \usepackage{lineno}
  - \usepackage{caption,setspace}
  - \captionsetup[figure]{font=small}
bibliography: ReferencesAbbrev
biblio-style: EcolLett
fontsize: 11pt
fontfamily: mathpazo
documentclass: article
linkcolor: black
urlcolor: black
citecolor: black
---

@@

@@

*Running title*: Character displacement destabilizes food webs

@@

<!-- Remove rest of title page for bioRxiv submission
\bigskip

*Manuscript type*: Letter.

*Number of words in abstract*: 143; 
*Number of words in main text*: `r wordcountaddin:::word_count("manuscript_ECD_model.Rmd") - 143`;
*Number of words in text boxes*: NA;
*Number of references*: 45;
*Number of figures*: 4;
*Number of tables*: 0;
*Number of text boxes*: 0

@@

\linenumbers{}
\modulolinenumbers[3]

@@

## color palette
cbPalette <- c("#D55E00", "#0072B2", "#009E73")
```

@@

To gain insight to the eco-evolutionary feedback generated by character displacement, I conducted simulations using an Adaptive Dynamics approach. 
Specifically, after letting consumer and resource abundances reach a steady state, I created a mutant consumer by randomly choosing one and modifying its attack rate on one resource by either subtracting or adding a small constant (0.01 in the following simulations) with equal probability. 
The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006]. 
This function has the useful property that it differentiates between cases where intermediate combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ are higher than the extremes (when $n>1$, green line in Fig. \ref{fig:tradeoff}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line in Fig. \ref{fig:tradeoff}). 
When $n=1$, the tradeoff function is linear, and all combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ have the same total attack rate (blue line in Fig. \ref{fig:tradeoff}). 
Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource abundances to reach a steady state. 
I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g., $\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}$ is constrained to a maximum of 1 and minimum of 0). 
Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and $w_{11} = w_{22} = 0.6$. 
I set an initial value of $a_{11} = a_{22} = 1.2$, while $a_{12}$ and $a_{21}$ depended on the value of *n*. 
I set initial consumer and resource abundances to: $R_1 = R_2 = 2$; $C_1 = C_2 = 1$.
All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model) and has been archived with Zenodo (https://zenodo.org/badge/latestdoi/79349192).

@@

tradeoff_plot <- bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
  ggplot(., aes(x=aii, y=aij, color = factor(n))) + 
  geom_line() + 
  geom_point(data = initial_points, aes(x=a11, y=a12), size=3, show.legend=FALSE) + 
  xlab(expression(bold(a[11]))) +
  ylab(expression(bold(a[21]))) + 
  scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette)

@@

```{r Fig 3, include=FALSE}
Macs <- plot_grid(
  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))) +
    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), 
  Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)), 
  Stability_Mac + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)), 
  ncol=1, align='v', labels = c("","b)","c)"), hjust = -4, vjust = 0.5, label_size = 10
)
Macs_title <- ggdraw() + draw_label("Resources: Same Habitat", fontface = 'bold', size = 10, hjust = 0.2)
plot_Macs <- plot_grid(Macs_title, Macs, ncol = 1, rel_heights = c(0.1, 1))

LSs <- plot_grid(
  LS_attack  + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])) + 
    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), 
  LS_resources+ xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank()),   
  Stability_LS + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab("") + theme(legend.position = "none", axis.text = element_text(size=10)), 
  ncol=1, align='v', labels = c("","e)","f)"), hjust = -1.5, vjust = 0.5, label_size = 10
)
LSs_title <- ggdraw() + draw_label("Resources: Different Habitat", fontface = 'bold', size = 10, hjust = 0.3)
plot_LSs <- plot_grid(LSs_title, LSs, ncol = 1, rel_heights = c(0.1, 1))

@@

# Acknowledgements

This work was inspired by discussions with Seth Rudman, Dolph Schluter, Ben Gilbert, and Kevin McCann.
Sally Otto provided much help in early analyses of the mathematical model. 
I would not have been able to take the theory as far as I did without her guidance and encouragement.
Matt Osmond, Jean Gibert, and Seth Rudman provided critical feedback on an earlier draft.
For funding support, I thank the University of British Columbia (Four-Year Fellowship to M.A. Barbour), NSERC (Discovery grant to Greg Crutsinger), and the Swiss National Science Foundation (grant 31003A_160671 to Jordi Bascompte).

>>>>

---
output:
  pdf_document:
    keep_tex: true
    toc: false
    number_sections: false
    df_print: paged
    fig_caption: true
    citation_package: natbib 
    #includes:
    #  after_body: Figures_Edited.tex
header-includes:
  - \usepackage{fullpage}
  - \linespread{1.7} 
  - \usepackage{lineno}
  - \usepackage{caption}
  - \captionsetup[figure]{font=small}
bibliography: ReferencesAbbrev.bib
biblio-style: amnat
fontsize: 11pt
fontfamily: mathpazo
documentclass: article
linkcolor: black
urlcolor: black
citecolor: black
---

@@

<!-- remove for double-blind peer-review
\bigskip

@@

<!--*Running title*: Character displacement destabilizes food webs-->

*Manuscript type*: E-Article.

*Manuscript elements*: Figures 1-4; Online appendices S1-4 (including figures S1-3 in appendix S4).

@@

@@

\linenumbers{}
\modulolinenumbers[2]

@@

## color palette
cbPalette <- c("#D55E00", "#0072B2", "#009E73")

## set plot theme
theme_set(theme_cowplot())
```

@@

To gain insight to the eco-evolutionary feedback generated by character displacement, I conducted simulations using an Adaptive Dynamics approach. 
Specifically, after letting consumer and resource abundances reach a steady state, I created a mutant consumer by randomly choosing one and modifying its attack rate on one resource by either subtracting or adding a small constant (0.01 in the following simulations) with equal probability. 
The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006]. 
This function has the useful property that it differentiates between cases where intermediate combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ are higher than the extremes (when $n>1$, green line in Fig. \ref{fig:tradeoff}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line in Fig. \ref{fig:tradeoff}). 
When $n=1$, the tradeoff function is linear, and all combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ have the same total attack rate (blue line in Fig. \ref{fig:tradeoff}). 
Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource abundances to reach a steady state. 
I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g., $\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}$ is constrained to a maximum of 1 and minimum of 0). 
Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and $w_{11} = w_{22} = 0.6$. 
I set an initial value of $a_{11} = a_{22} = 1.2$, while $a_{12}$ and $a_{21}$ depended on the value of *n*. 
I set initial consumer and resource abundances to: $R_1 = R_2 = 2$; $C_1 = C_2 = 1$.
All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub <!--(https://github.com/mabarbour/ECD_model) -->and has been archived with Zenodo <!-- (https://zenodo.org/badge/latestdoi/79349192)-->(please contact journal office for links).

@@

tradeoff_plot <- bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
  ggplot(., aes(x=aii, y=aij, color = factor(n))) + 
  geom_line() + 
  geom_point(data = initial_points, aes(x=a11, y=a12), size=3, show.legend=FALSE) + 
  xlab(expression(bold(a[11]))) +
  ylab(expression(bold(a[21]))) + 
  scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette) + theme_cowplot()

@@

```{r Fig 3, include=FALSE}
Macs <- plot_grid(
  Mac_attack + xlab("") + ylab(bquote(atop("Effective attack rate","("*a[1*","*j]+a[2*","*j]*")"))) +
    theme_cowplot() +
    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(vjust=0, size=12)), 
  Mac_resources + xlab("") + ylab(bquote(atop("Resource abundance","("*R[1]+R[2]*")"))) + 
    theme_cowplot() +
    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), axis.title.y=element_text(vjust=0, size=12)), 
  Stability_Mac + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab(bquote(atop("Food-web stability","("*-lambda[max]*")"))) + 
    theme_cowplot() +
    theme(legend.position = "none", axis.text = element_text(size=10), axis.title.y=element_text(vjust=0, size=12)), 
  ncol=1, align='v', labels = c("","b)","c)"), hjust = -4, vjust = 0.5, label_size = 10
)
Macs_title <- ggdraw() + draw_label("Resources: Same Habitat", fontface = 'bold', size = 10, hjust = 0.2)
plot_Macs <- plot_grid(Macs_title, Macs, ncol = 1, rel_heights = c(0.1, 1))

LSs <- plot_grid(
  LS_attack  + xlab("") + ylab(bquote(w[1*","*j]*a[1*","*j]+w[2*","*j]*a[2*","*j])) + 
    theme_cowplot() +
    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank(), plot.title = element_text(size = 10), axis.title.y=element_text(size=12)), 
  LS_resources + xlab("") + ylab("") +
    theme_cowplot() +
    theme(legend.position = "none", axis.text = element_text(size=10), axis.text.x = element_blank()),   
  Stability_LS + scale_x_continuous(labels = c(0,0.25,0.5,0.75,1)) + xlab("") + ylab("") + 
    theme_cowplot() +
    theme(legend.position = "none", axis.text = element_text(size=10)), 
  ncol=1, align='v', labels = c("","e)","f)"), hjust = -1.5, vjust = 0.5, label_size = 10
)
LSs_title <- ggdraw() + draw_label("Resources: Different Habitat", fontface = 'bold', size = 10, hjust = 0.3)
plot_LSs <- plot_grid(LSs_title, LSs, ncol = 1, rel_heights = c(0.1, 1))

@@

<!-- Remove for double-blind review
# Acknowledgements

This work was inspired by discussions with Seth Rudman, Dolph Schluter, Ben Gilbert, and Kevin McCann.
Sally Otto provided much help in early analyses of the mathematical model. 
I would not have been able to take the theory as far as I did without her guidance and encouragement.
Matt Osmond, Jean Gibert, and Seth Rudman provided critical feedback on an earlier draft.
For funding support, I thank the University of British Columbia (Four-Year Fellowship to M.A. Barbour), NSERC (Discovery grant to Greg Crutsinger), and the Swiss National Science Foundation (grant 31003A_160671 to Jordi Bascompte).
-->

### 2020-01-30 manuscript_ECD_model.tex
omitted personal identifying information for double-blind peer review at Am Nat

\documentclass[11pt,]{article}
\usepackage[]{mathpazo}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=black,
            urlcolor=black,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{EcolLett}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{}
    \pretitle{\vspace{\droptitle}}
  \posttitle{}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{fullpage}
\linespread{1.5}
\usepackage{lineno}
\usepackage{caption,setspace}
\captionsetup[figure]{font=small}

@@

\begin{center} \large Matthew A. Barbour$^{1,2,\ast}$ \normalsize \end{center}

@@

\noindent 1. University of Zurich, Department of Evolutionary Biology
and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich,
Switzerland;

\noindent 2. University of British Columbia, Department of Zoology, 2212
Main Mall, Vancouver, BC V6T 1Z4, Canada;

\(^\ast\) Corresponding author; e-mail:
\href{mailto:matthew.barbour@ieu.uzh.ch}{\nolinkurl{matthew.barbour@ieu.uzh.ch}},
phone: +41 44 635 61 63

\emph{Statement of authorship}: MAB conceived the study, derived
analytical solutions, performed simulations, and wrote the paper.

\emph{Data accessibility}: No new data were used. All code used to
reproduce this manuscript and simulations within are publically
available on GitHub (\url{https://github.com/mabarbour/ECD_model}).

\emph{Running title}: Character displacement destabilizes food webs

@@

To gain insight to the eco-evolutionary feedback generated by character
displacement, I conducted simulations using an Adaptive Dynamics
approach. Specifically, after letting consumer and resource abundances
reach a steady state, I created a mutant consumer by randomly choosing
one and modifying its attack rate on one resource by either subtracting
or adding a small constant (0.01 in the following simulations) with
equal probability. The mutant's attack rate on the other resource was
determined by a tradeoff, such that
\(\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1\),
where \(A\) is the total investment in attack rates and \(n\) describes
the shape of the tradeoff \citep{Sargent2006}. This function has the
useful property that it differentiates between cases where intermediate
combinations of \(a_{\text{1},j}\) and \(a_{\text{2},j}\) are higher
than the extremes (when \(n>1\), green line in Fig. \ref{fig:tradeoff})
or, conversely, where the two extremes are higher than intermediate
investments (when \(n<1\), orange line in Fig. \ref{fig:tradeoff}). When
\(n=1\), the tradeoff function is linear, and all combinations of
\(a_{\text{1},j}\) and \(a_{\text{2},j}\) have the same total attack
rate (blue line in Fig. \ref{fig:tradeoff}). Assuming the mutant
consumer was rare, I then determined whether the mutant had higher
relative fitness than the resident consumer, and thus could invade and
replace the resident consumer. If the mutant was able to invade, I
updated the attack rate of the resident consumer to the mutant attack
rate and allowed consumer and resource abundances to reach a steady
state. I then repeated the simulation up to 10,000 times, which was
sufficient for consumers to either reach an evolutionary stable strategy
\citep[ESS,][]{Smith1973} or an evolutionary limit (e.g.,
\(\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}\) is constrained to a
maximum of 1 and minimum of 0). Unless otherwise noted, I conducted
simulations with the following parameter values: \emph{r} = 1; \emph{K}
= 4; \emph{e} = 0.8; \emph{m} = 1; \emph{A} = 2; \emph{h} = 0.4; and
\(w_{11} = w_{22} = 0.6\). I set an initial value of
\(a_{11} = a_{22} = 1.2\), while \(a_{12}\) and \(a_{21}\) depended on
the value of \emph{n}. I set initial consumer and resource abundances
to: \(R_1 = R_2 = 2\); \(C_1 = C_2 = 1\). All simulations were conducted
in R \citep{R} and the code to reproduce these simulations is publically
available on GitHub (\url{https://github.com/mabarbour/ECD_model}).

@@

\section{Acknowledgements}\label{acknowledgements}

This work was inspired by discussions with Seth Rudman, Dolph Schluter,
Ben Gilbert, and Kevin McCann. Sally Otto provided much help in early
analyses of the mathematical model. I would not have been able to take
the theory as far as I did without her guidance and encouragement. Matt
Osmond, Jean Gibert, and Seth Rudman provided critical feedback on an
earlier draft. For funding support, I thank the University of British
Columbia (Four-Year Fellowship to M.A.~Barbour), NSERC (Discovery grant
to Greg Crutsinger), and the Swiss National Science Foundation (grant
31003A\_160671 to Jordi Bascompte).

\bibliography{ReferencesAbbrev}

>>>>

\documentclass[11pt,]{article}
\usepackage[]{mathpazo}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            colorlinks=true,
            linkcolor=black,
            citecolor=black,
            urlcolor=black,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{amnat}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{fullpage}
\linespread{1.7}
\usepackage{lineno}
\usepackage{caption}
\captionsetup[figure]{font=small}

\author{}
\date{\vspace{-2.5em}}

@@

\emph{Manuscript type}: E-Article.

@@

\emph{Manuscript elements}: Figures 1-4; Online appendices S1-4
(including figures S1-3 in appendix S4).

@@

To gain insight to the eco-evolutionary feedback generated by character
displacement, I conducted simulations using an Adaptive Dynamics
approach. Specifically, after letting consumer and resource abundances
reach a steady state, I created a mutant consumer by randomly choosing
one and modifying its attack rate on one resource by either subtracting
or adding a small constant (0.01 in the following simulations) with
equal probability. The mutant's attack rate on the other resource was
determined by a tradeoff, such that
\(\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1\),
where \(A\) is the total investment in attack rates and \(n\) describes
the shape of the tradeoff \citep{Sargent2006}. This function has the
useful property that it differentiates between cases where intermediate
combinations of \(a_{\text{1},j}\) and \(a_{\text{2},j}\) are higher
than the extremes (when \(n>1\), green line in Fig. \ref{fig:tradeoff})
or, conversely, where the two extremes are higher than intermediate
investments (when \(n<1\), orange line in Fig. \ref{fig:tradeoff}). When
\(n=1\), the tradeoff function is linear, and all combinations of
\(a_{\text{1},j}\) and \(a_{\text{2},j}\) have the same total attack
rate (blue line in Fig. \ref{fig:tradeoff}). Assuming the mutant
consumer was rare, I then determined whether the mutant had higher
relative fitness than the resident consumer, and thus could invade and
replace the resident consumer. If the mutant was able to invade, I
updated the attack rate of the resident consumer to the mutant attack
rate and allowed consumer and resource abundances to reach a steady
state. I then repeated the simulation up to 10,000 times, which was
sufficient for consumers to either reach an evolutionary stable strategy
\citep[ESS,][]{Smith1973} or an evolutionary limit (e.g.,
\(\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}\) is constrained to a
maximum of 1 and minimum of 0). Unless otherwise noted, I conducted
simulations with the following parameter values: \emph{r} = 1; \emph{K}
= 4; \emph{e} = 0.8; \emph{m} = 1; \emph{A} = 2; \emph{h} = 0.4; and
\(w_{11} = w_{22} = 0.6\). I set an initial value of
\(a_{11} = a_{22} = 1.2\), while \(a_{12}\) and \(a_{21}\) depended on
the value of \emph{n}. I set initial consumer and resource abundances
to: \(R_1 = R_2 = 2\); \(C_1 = C_2 = 1\). All simulations were conducted
in R \citep{R} and the code to reproduce these simulations is publically
available on GitHub and has been archived with Zenodo (please contact
journal office for links).

@@

\bibliography{ReferencesAbbrev.bib}
